---
title: "Explore"
author: "Yogi Patel"
date: "October 22, 2015"
output: html_document
---


```{r}
require(ggplot2)
require(grid)
require(plyr)
data(diamonds)
data(mtcars)

# Recreating Diamonds$VS column
# Determining the ratio of 1's and 0's in mtcars$vs
x <- count(mtcars, 'vs')
# Calculating ratio of 1's in dataframe
propOnes <- (x$freq[2]/(x$freq[2]+x$freq[1]))
# Determining number of rows in diamonds dataframe
numObs <- nrow(diamonds)
# Calculating number of Trues's needed in diamonds dataframe
numTrue <- propOnes * numObs
# Getting a whole number value for numTrue
numTrue <- round(numTrue,0)
# Getting number of false
numFalse <- numObs - numTrue
# Creating 'vs' column in diamonds
diamonds$vs <- as.logical(rep(c(0,1), c(numTrue, numFalse)))

```

## Creating Explore Function
```{r}

explore <- function(df, vec, thresh) {
  
  # Question One:
  # Loading ggplot
  library(ggplot2)
  
  # Getting the numeric columns 
  numCol <- df[which(lapply(df, is.numeric) == TRUE)]

  # Iterating through numeric columns
  for (i in 1:length(numCol)) {
  # Iterate through bin sizes
    for (k in vec) {
      # Creating a histogram with counts
      # Initializing plot
      count <- ggplot(numCol, aes(x = numCol[[i]]), environment=environment())
      # Setting plot to histogram, with blue as the bin fill color and "k" as the width
      count <- count + geom_histogram(fill = 'blue', binwidth = k) +
               # Adding X, Y, and Title labels to the graph
               labs(x = names(numCol[i]), y = "Count", title = paste("Bin size:", k)) +
               # Adding a vertical line at the mean
               geom_vline(xintercept = colMeans(numCol[i]), colour='red')
      # Updating the aesthetic of count graph to get a density histrogram
      density <- count + aes(y = ..density..) + labs(y = "Density")
      # Printing the histograms
      print(count)
      print(density)    
    } 
  } 
  
  # Question Two:
  # Getting the categorical and binary variables
  catCol <- cbind(df[which(lapply(df, is.logical) == TRUE)], df[which(lapply(df, is.factor) == TRUE)])
  
  # Iterating through all categorical columns
  for (i in 1:length(catCol)) {
    # Initializing plot
    gbar <- ggplot(catCol, aes(x = catCol[[i]]), environment=environment())
    # Generating grey bar chart and adding x and y labels
    gbar <- gbar + geom_bar(fill = 'grey') + labs(x = names(catCol[i]), y = "Count")
    # Printing the bar graph
    print(gbar)
  }
  
  # Question Three:
  
  # Creating Pearson's correlation coefficient table
  cors <- cor(numCol, method = "pearson")
  
  # Creating a variable for row count
  row <- 0
  # Creating empty vectors for pairs and rsquare values
  pairs <- c()
  rsquare <- c()
  
  # Iterating through columns of the correlation table
  for (i in 1:(ncol(cors)-1)) {
    # Iterating through next columns of correlation table
    for (k in (i+1):ncol(cors)){
      # Incresing vector row count
      row <- row + 1
      # Adding name of variables to pair list
      pairs[row] <- paste(colnames(cors)[i], colnames(cors)[k], sep="-")
      # Adding r-sqaure value for pair of variables to r-square vector
      rsquare[row] <- cors[i,k]^2
    }
  }
  
  # Creating a dataframe from the name combinations and rsquare vectors
  rsquares <- data.frame(pairs, rsquare)
  # Updating column names for the rsquares data frame
  names(rsquares) <- c('Pairs', "R-Squared")
  
  
  # Question 4
  
  # Intiating an empty object that will be the list we ultimately return from function
  returnList <- NULL
  
  
  # Creating frequency tables for cateogorical and binary variables
  for (i in 1:length(catCol)) {
    # Creating frequecny table for the column
    freq <- table(catCol[i])
    # Converting Frequency table into dataframe
    freq <- data.frame(freq)
    # Updating Column names to include variable name
    names(freq) <- c(names(catCol)[i], "Frequency")
    # Adding frequency table to the list
    returnList <- c(returnList, list(freq))
  }
  
  # Adding the R-Sqaures dataframe (from Q3) to returnList
  returnList <- c(returnList, list(rsquares))
  
  
  # Initializing variables to be used in nested loop below
  row <- 0   # Creating a variable that will hold the number of rows in the output dataframe
  pairs <- c() # Creating an empty vector to hold the pairs of variables
  corrs <- c() # Creating an empty vector to hold the correlations

  # Iterating through the variables in the new (numeric only) data frame, starting with the first
  # variable and ending with the second to last variable.
  for( a in 1:(ncol(numCol)-1)) {
    # Now interating through each of the subsequent variables left in the dataframe.
    for(b in (a+1):(ncol(numCol))) {
      # Checking if Pearson Coefficient is larger than threshold
      if (cor(numCol[,a], numCol[,b], method="pearson") > thresh) {
         # Adding a row to the dataframe
      row <- row + 1
      # Adding the names of the columns to the pairs vector
      pairs[row] <- paste(colnames(numCol)[a], colnames(numCol)[b], sep="-")
      # Calculating the Pearson Coefficient of the row
      corrs[row] <- cor(numCol[,a], numCol[,b], method="pearson")
    }
      }
     
  }

  # Create a new dataframe. There will be 2 columns, one for pairs and one for corrs.
  threshCorrs <- data.frame(pairs, corrs)
  # Updating the column names
  names(threshCorrs) <- c('Pairs', "Pearsons's Corr.")
  
  # Adding threshCorrs to returnList - any correlation greater than our threshold.
  returnList <- c(returnList, list(threshCorrs))s
  
  
  # Returning final list for function
  return(returnList)

  }
  
# Question 5:
# Testing the function explore

explore(diamonds, list(5, 20, 50), .25)  


# Looks like it's working to me. 
