---
title: "Homework 4"
output: html_document
---
**by Yogi Patel**


**Question One:**

*A.*  Creating a vector from 1 - 20.
```{r}
vec1 <- c(1:20)
vec1
```

*B.*  Creating a vector that reverses `vec1`
```{r}
vec2 <- rev(vec1)
vec2
```

*C.*  Combining the vectors `vec1` and `vec`
```{r}
vec3 <- c(vec1[1:19],vec2)
vec3
```

*D.*  Creating a vector named `tmp`
```{r}
tmp <- c(4,6,3)
tmp
```

*E.*  Replicating `tmp` 10 times
```{r}
tmp2 <- rep(tmp, 10)
tmp2
```

*F.*  Adding a 4 to the end of `tmp3`
```{r}
tmp3 <- c(tmp2, 4)
tmp3
```

*G.*  Creating vector `tmp4`
```{r}
tmp4 <- rep(tmp, c(10,20,30))
tmp4
```


**Question Two**

Creating a vector of exp(x)*cos(x) for x=(3,3.1,3.2,...,5.9,6)
```{r}
x <- seq(3, 6, by=0.1)
x
expCos <- exp(x)*cos(x)
expCos
```

**Question Three**

*A.* 
Creating vector for exponents for `.1`
```{r}
a <- seq(3, 36, by=3)
```
Creating a vector of `.1` the length of `a`
```{r}
b <- rep(.1, length(a))
```
Creating a vector that exponentiates vector `b` by vector `a`
```{r}
c <- b^a
```
Creating vector of exponents for `.2`
```{r}
d <- seq(1, 34, 3)
```
Creating a vector of `.2` the length of `d`
```{r}
e <- rep(.2, length(d))
```
Creating a vectore that exponentiates vector `e` by vector `d`
```{r}
f <- e^d
```
Multiplying the vectors `c` and `f` for final answer
```{r}
q3A <- c *f
q3A
```

*B.*
Creating the denominator and exponent vector `denEx`
```{r}
denEx <- c(1:25)
```
Creating the numerator vector `numer` which is a sequnce of 2 in length 25
```{r}
numer <- rep(2, 25)
```
Creating the final vector
```{r}
q3B <- (numer^denEx)/denEx
q3B
```

**Question Four**

*A.*

Creating initial sequnce of a all `i` values.
```{r}
i <- c(10:100)
```
Performing calculation for each `i` value.
```{r}
iCalc <- ((i^3) + (4*(i^2)))
```
Performing the summation of `iCalc`.
```{r}
sum1 <- sum(iCalc)
sum1
```


*B.*
Creating the initial sequence of `i` values.
```{r}
i <- c(1:25)
```
Performing calculation for each `i` value.
```{r}
iCalc <- (((2^i)/i) + ((3^i)/(i^2)))
```
Performing the summation of `iCalc`.
```{r}
sum2 <- sum(iCalc)
sum2
```


** Question Five **

*A.* Creating a vector of 30 labels, with a space.
```{r}
labels <- paste("label", 1:30, sep = " ")
labels
```

*B.* Creating a vector of 30 fns, without a space.
```{r}
fns <- paste("fn", 1:30, sep = "")
fns
# SW: see also paste0()
```


** Question Six **
Creating two random vectors `xVec` and `yVec`
```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
xVec
yVec
# SW: lost points: you needed to use set.seed() to ensure reproducible results; 
# I've added it for you for the remaining problems
```

*A.* Creating a vector of `yVec(i) - xVec(i-2)`
```{r}
newVec1 <- (yVec[2:250] - xVec[1:249])
newVec1
# Checking length... should be 249
length(newVec1)
```

*B.* Creating a vector of `sin(y[i-1])/cos(x[1])`
```{r}
newVec2 <- (sin(yVec[1:249])/cos(xVec[2:250]))
newVec2
```

*C.* Creating a vector of `x(1)+2x(2)-x(3)`
```{r}
newVec3 <- xVec[1:248]+(2*xVec[2:249])-xVec[3:250]
newVec3
# Checking length... should be 248
length(newVec3)
```

*D.* 

Creating a vector for the summation
```{r}
sumVec <- (exp((xVec[2:250]*-1)))/(xVec[1:249]+10)
sumVec
# Checking length... should be 249
length(sumVec)
```

Performing the summation of `sumVec`
```{r}
summation <- sum(sumVec)
summation
```


**Question Seven**
*A.* Picking out values of `yVec` that are greater than 600 and storing as `y600`
```{r}
y600 <- yVec[yVec > 600]
y600
```

*B.* Getting index positions of values greater than 600 in `yVec`
```{r}
y600Index <- which(yVec > 600)
y600Index

# As a check, just making sure the vector or values and vector of indexes are the same length
length(y600) == length(y600Index)

```

*C.* Getting values of `xVec` that correspond to the indexes of where `yVec` is greater than 600.
```{r}
xIndex <- xVec[y600Index]
xIndex
```

*D* Creating a vector of `sqrt(abs(x - xbar))`
```{r}
q7D <- sqrt(abs(xVec - mean(xVec)))
q7D
```

*E* Calculating number of values in `yVec` that are within 200 of the max value of vVec
```{r}
ymax200 <- length(yVec[yVec <= (max(yVec)-200)])
ymax200

# 194 values in yVec are within 200 of the max value
# SW: lost points: look at the values of ymax200, many of them clearly are not 
# within 200 of max(yVec). your code is subtracting 200 from the max, then 
# returning all values of yVec less than that. that's a different question than 
# all values _within_ 200 of the max. here's the code you need:
length(yVec[abs(yVec - max(yVec)) <= 200])
```


*F.* Calculating number of values in `xVec` that are divisible by 2
```{r}
evenX <- length(xVec[(xVec %% 2) == 0 ])
evenX

# 129 values in xVec are divisible by 2
```


*G.* Sorting `xVec` by order of `yVec`
```{r}
sort(xVec, partial = order(yVec))
xVec
# SW: lost points: incorrect sort, you want
xVec[order(yVec)]
```

*H.*

Creating the vector of indeces
```{r}
indexes <- seq(1, 250, by=3)
indexes
```

Creating a vector picking on values in `yVec` from `indexes`
```{r}
newY <- yVec[indexes]
newY

length(indexes) == length(newY)
```

** Question 8 **
Creating the `numerator` vector
```{r}
numerator <- seq(2,38,by=2)
```

Creating the `denominator` vector
```{r}
denominator <- seq(3,39,by=2)
```

Dividing `numerator` by `denominator`
```{r}
numdem <- numerator / denominator
```

Calculating the cumulative product
```{r}
prodVec <- cumprod(numdem)
```

Adding 1 to the summation of `prodVec` for the final answer
```{r}
answer8 <- 1 + sum(prodVec)
answer8
```

