---
title: "HW5"
output: html_document
---

SW: nice job.

```{r}
require(ggplot2)
require(grid)
data(diamonds)

```

**Question 1**

Using `sapply` to extract all numeric columns from the data frame.

The fuction `numericCols` takes one input parameter, a dataframe `df`. The function returns a dataframe of only the numeric variables contained in the input parameter.
```{r}
numericCols <- function(df){
   dfNum <- sapply(df, is.numeric)
   df <- df[dfNum]
   return(df)
}

# Testing that Function only selects numeric cols
head(numericCols(diamonds))
```

**Question 2**

The function `columnCorr` takes one input parameters, a data frame `df`. The function returns a dataframe of the Pearson correlation coeffcients for each pair of numeric columns, making the use of function `numericCols` defined above. 
```{r}
columnCorr <- function(df){
  # Using numericCols to get a dataframe of just numeric columns
  df <- numericCols(df)
  
  # Initializing variables to be used in nested loop below
  row <- 0   # Creating a variable that will hold the number of rows in the output dataframe
  pairs <- c() # Creating an empty vector to hold the pairs of variables
  corrs <- c() # Creating an empty vector to hold the correlations
  
  # Iterating through the variables in the new (numeric only) data frame, starting with the first
  # variable and ending with the second to last variable.
  for( a in 1:(ncol(df)-1)) {
    # Now interating through each of the subsequent variables left in the dataframe.
    for(b in (a+1):(ncol(df))) {
      # Adding a row to the dataframe
      row <- row + 1
      # Adding the names of the columns to the pairs vector
      pairs[row] <- paste(colnames(df)[a], colnames(df)[b], sep="-")
      # Calculating the Pearson Coefficient of the row
      corrs[row] <- cor(df[,a], df[,b], method="pearson")
    }
  }
  
  # Create a new dataframe. There will be 2 columns, one for pairs and one for corrs.
  df <- data.frame(pairs, corrs)
  # Updating the column names
  names(df) <- c('Pairs', "Pearsons's Corr.")
  return(df) 
}

columnCorr(diamonds)
```

**Question 3**

```{r}

df <- numericCols(diamonds)
  
# Iterating through the variables in the new (numeric only) data frame, starting with the first
# variable and ending with the second to last variable.
for( a in 1:(ncol(df)-1)) {
  # Now interating through each of the subsequent variables left in the dataframe.
  for(b in (a+1):(ncol(df))) {
    # Data to plot
    data <- data.frame(df[,a],df[,b])
    # Adding axis lables to plotting data
    x <- colnames(df)[a]
    y <- colnames(df)[b]
    colnames(data) <- c(x,y)
    # Calculating Pearson Corr for the title
    corr <- cor(df[,a], df[,b], method="pearson")
    
    plots <- ggplot(data, aes_string(x=x,y=y)) + geom_point() + ggtitle(paste("Pearson Correlation for ",x,"and",y,":",corr))
    plot(plots)
    }
  }
  

```