---
title: "TEST2_R"
author: "Yogi Patel"
date: "November 26, 2015"
output: html_document
---

Loading the abalone dataset

```{r}
uciaddress <- "http://archive.ics.uci.edu/ml/machine-learning-databases/"
dataset <- "abalone/abalone.data"
getdataset <- paste(uciaddress, dataset, sep="")
abalone <- read.csv(getdataset)
str(abalone)

# Naming the columns
colnames(abalone) <- c("Gender","Length","Diameter",
                       "Height", "Whole_wgt",
                       "Shucked_wgt", "Viscera wgt",
                       "Shell wgt", "Rings")
str(abalone)
```

####Question 1

```{r}

# To get random samples, using the random samples function from class. 

# Declaring functions
randomSample = function(dataframe,n,replacement) {     
  return (dataframe[sample(nrow(dataframe), n, replace=replacement),]) }

# Calculating sample sizes for 1%, 10%, 25%, 50%, 80% samples
n1 <- as.integer(.01 * nrow(abalone))
n10 <- as.integer(.10 * nrow(abalone))
n25 <- as.integer(.25 * nrow(abalone))
n50 <- as.integer(.5 * nrow(abalone))
n80 <- as.integer(.8 * nrow(abalone))

# Getting random samples
abSample1 <- randomSample(abalone, n1, FALSE)
abSample10 <- randomSample(abalone, n10, FALSE)
abSample25 <- randomSample(abalone, n25, FALSE)
abSample50 <- randomSample(abalone, n50, FALSE)
abSample80 <- randomSample(abalone, n80, FALSE)

# Creating dataframes with min, max, mean, and variance for each sample 
abSample1 <- data.frame(c( min(abSample1$Whole_wgt), max(abSample1$Whole_wgt), mean(abSample1$Whole_wgt), var(abSample1$Whole_wgt) ))
abSample10 <- data.frame(c( min(abSample10$Whole_wgt), max(abSample10$Whole_wgt), mean(abSample10$Whole_wgt), var(abSample10$Whole_wgt) ))
abSample25 <- data.frame(c( min(abSample25$Whole_wgt), max(abSample25$Whole_wgt), mean(abSample25$Whole_wgt), var(abSample25$Whole_wgt) ))
abSample50 <- data.frame(c( min(abSample50$Whole_wgt), max(abSample50$Whole_wgt), mean(abSample50$Whole_wgt), var(abSample50$Whole_wgt) ))
abSample80 <- data.frame(c( min(abSample80$Whole_wgt), max(abSample80$Whole_wgt), mean(abSample80$Whole_wgt), var(abSample80$Whole_wgt) ))
abSample100 <- data.frame(c( min(abalone$Whole_wgt), max(abalone$Whole_wgt), mean(abalone$Whole_wgt), var(abalone$Whole_wgt) ))

# Combining dataframes of summary statistics from samples
SampleStats <- cbind(abSample1, abSample10, abSample25, abSample50, abSample80, abSample100)

# Naming rows and columns of SampleStats dataframe
colnames(SampleStats) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample", "100% Sample")
rownames(SampleStats) <- c("SampleMin", "SampleMax", "SampleMean", "SampleVariance")

# 6 columns and 4 fours, just like the question asks.
str(SampleStats)

```


####Question 2

```{r}
# Loading necessary libraries
library(ggplot2)
library(grid)

# Creating function creates a grid of histograms for each row of the input dataframe
histGrid <- function(df, transpose=FALSE) {
  
  # If transpose = true, transposing the the input dataframe before creating graphs
  if (transpose == TRUE) {
    df <- data.frame(t(df))
  }
  
  # Getting number of columns the grid will need
  cols <- round(ncol(df)/2)
  
  # Creating a new grid page
  grid.newpage()
  
  # Creating layout for grid 
  pushViewport(viewport(layout = grid.layout(cols, 2)))
  
  # Creating vplayout function to help place charts
  vplayout <- function(x,y,a){ 
    viewport(layout.pos.row = x, layout.pos.col = y,angle = a) 
  }
  
  # Looping through first half of the statistics
  for (k in 1:cols) {
    # Creating variable to store column name
    x.stat <-colnames(df)[k]
    # Creating an empty plot
    hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=df, aes_string(x=x.stat, y ="..density.."), binwidth=0.01)
    # Pushing the plot to the grid using vplayout function created above
    print(hist, vp = vplayout(k,1,0))
  }
   # Looping through second half of the statistics 
  for (k in (cols+1):ncol(df)) {
    # Creating variable to store column name
    x.stat <-colnames(df)[k]
    # Creating an empty plot
    hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=df, aes_string(x=x.stat, y ="..density.."), binwidth=0.01)
    # Pushing the plot to the grid using vplayout function created above
    print(hist, vp = vplayout((k-cols),2,0))
  }

}

# Running function of 
histGrid(SampleStats, transpose=TRUE)

```


####Question 3

```{r}

# Creating an empty dataframe that will be used to store 1000 samples
Samples1000 <- data.frame("1% Sample" = 0, "10% Sample" = 0, "25% Sample" = 0, "50% Sample" = 0 , "80% Sample" = 0)

# Starting a for loop that iterates 1000 times
for(i in 1:1000) {
  
  # Taking random sample using sample size determined above 
  # Getting random samples
  abSample1 <- randomSample(abalone, n1, FALSE)
  abSample10 <- randomSample(abalone, n10, FALSE)
  abSample25 <- randomSample(abalone, n25, FALSE)
  abSample50 <- randomSample(abalone, n50, FALSE)
  abSample80 <- randomSample(abalone, n80, FALSE)
  
  # Adding sample to the empty dataset
  Samples1000[i,]<- c(mean(abSample1$Whole_wgt),
                      mean(abSample10$Whole_wgt),
                      mean(abSample25$Whole_wgt),
                      mean(abSample50$Whole_wgt),
                      mean(abSample80$Whole_wgt))
  
}

# Now the dataframe has four columns coressponding to each the four calculated sample statistcs and 1000 rows for 1000 samples.
# The 1%, 10%, 25%, 50%, and 80% sample sizes each make up 1/5 of the dataframe
str(Samples1000)

```


###Question 4

```{r}

# Adding the 100% sample to the dataframe
Samples1000[,6] <- abSample100[3,]

# Naming rows and columns of Samples1000 dataframe
colnames(Samples1000) <- c("x1_Sample", "x10_Sample", "x25_Sample", "x50_Sample", 
                           "x80_Sample", "x100_Sample")

# Creating grid of histograms
histGrid(Samples1000)

```


###Question 5

```{r}

# Adding Mins Samples1000 to SampleStats
SampleStats[5,] <- c(min(Samples1000[,1]), max(Samples1000[,2]), min(Samples1000[,3]), 
                     min(Samples1000[,4]), min(Samples1000[,5]), min(Samples1000[,6]))

# Adding Maxes Samples1000 to SampleStats
SampleStats[6,] <- c(max(Samples1000[,1]), max(Samples1000[,2]), max(Samples1000[,3]), 
                     max(Samples1000[,4]), max(Samples1000[,5]), max(Samples1000[,6]))

# Adding Means Samples1000 to SampleStats
SampleStats[7,] <- c(mean(Samples1000[,1]), mean(Samples1000[,2]), mean(Samples1000[,3]), 
                     mean(Samples1000[,4]), mean(Samples1000[,5]), mean(Samples1000[,6]))

# Adding variances Samples1000 to SampleStats
SampleStats[8,] <- c(var(Samples1000[,1]), var(Samples1000[,2]), var(Samples1000[,3]), 
                     var(Samples1000[,4]), var(Samples1000[,5]), var(Samples1000[,6]))

# Updating row names for newly added columns
rownames(SampleStats)[5:8] <- c("ResampleMin", "ResampleMax", "ResampleMean", "ResampleVariance")

# New SampleStats dataframe - transposed with new Resample rows added
SampleStats
```


###Question 6
```{r}
# Creating a histogram
sampleMeans <- data.frame(t(SampleStats["SampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=abalone, aes(x=Whole_wgt, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =sampleMeans,(aes(xintercept = SampleMean, colour = "red" )))
print(hist)
```


###Question 7
```{r}
# Creating a histogram
resampleMeans <- data.frame(t(SampleStats["ResampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=abalone, aes(x=Whole_wgt, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =resampleMeans,(aes(xintercept = ResampleMean, colour = "red" )))
print(hist)
```

##Question 8 --- START

####Question 1 `diamonds$carat`

```{r}

# To get random samples, using the random samples function from class. 

# Calculating sample sizes for 1%, 10%, 25%, 50%, 80% samples
n1 <- as.integer(.01 * nrow(diamonds))
n10 <- as.integer(.10 * nrow(diamonds))
n25 <- as.integer(.25 * nrow(diamonds))
n50 <- as.integer(.5 * nrow(diamonds))
n80 <- as.integer(.8 * nrow(diamonds))

# Getting random samples
abSample1 <- randomSample(diamonds, n1, FALSE)
abSample10 <- randomSample(diamonds, n10, FALSE)
abSample25 <- randomSample(diamonds, n25, FALSE)
abSample50 <- randomSample(diamonds, n50, FALSE)
abSample80 <- randomSample(diamonds, n80, FALSE)

# Creating dataframes with min, max, mean, and variance for each sample 
abSample1 <- data.frame(c( min(abSample1$carat), max(abSample1$carat), mean(abSample1$carat), var(abSample1$carat) ))
abSample10 <- data.frame(c( min(abSample10$carat), max(abSample10$carat), mean(abSample10$carat), var(abSample10$carat) ))
abSample25 <- data.frame(c( min(abSample25$carat), max(abSample25$carat), mean(abSample25$carat), var(abSample25$carat) ))
abSample50 <- data.frame(c( min(abSample50$carat), max(abSample50$carat), mean(abSample50$carat), var(abSample50$carat) ))
abSample80 <- data.frame(c( min(abSample80$carat), max(abSample80$carat), mean(abSample80$carat), var(abSample80$carat) ))
abSample100 <- data.frame(c( min(diamonds$carat), max(diamonds$carat), mean(diamonds$carat), var(diamonds$carat) ))

# Combining dataframes of summary statistics from samples
SampleStats <- cbind(abSample1, abSample10, abSample25, abSample50, abSample80, abSample100)

# Naming rows and columns of SampleStats dataframe
colnames(SampleStats) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample", "100% Sample")
rownames(SampleStats) <- c("SampleMin", "SampleMax", "SampleMean", "SampleVariance")

# 6 columns and 4 fours, just like the question asks.
str(SampleStats)

```


####Question 2 `diamonds$carat`

```{r}

# Running function of 
histGrid(SampleStats, transpose=TRUE)

```


####Question 3 `diamonds$carat`

```{r}

# Creating an empty dataframe that will be used to store 1000 samples
Samples1000 <- data.frame("1% Sample" = 0, "10% Sample" = 0, "25% Sample" = 0, "50% Sample" = 0 , "80% Sample" = 0)

# Starting a for loop that iterates 1000 times
for(i in 1:1000) {
  
  # Taking random sample using sample size determined above 
  # Getting random samples
  abSample1 <- randomSample(diamonds, n1, FALSE)
  abSample10 <- randomSample(diamonds, n10, FALSE)
  abSample25 <- randomSample(diamonds, n25, FALSE)
  abSample50 <- randomSample(diamonds, n50, FALSE)
  abSample80 <- randomSample(diamonds, n80, FALSE)
  
  # Adding sample to the empty dataset
  Samples1000[i,]<- c(mean(abSample1$carat),
                      mean(abSample10$carat),
                      mean(abSample25$carat),
                      mean(abSample50$carat),
                      mean(abSample80$carat))
  
}

# Now the dataframe has four columns coressponding to each the four calculated sample statistcs and 1000 rows for 1000 samples.
# The 1%, 10%, 25%, 50%, and 80% sample sizes each make up 1/5 of the dataframe
str(Samples1000)

```


###Question 4 `diamonds$carat`

```{r}

# Adding the 100% sample to the dataframe
Samples1000[,6] <- abSample100[3,]

# Naming rows and columns of Samples1000 dataframe
colnames(Samples1000) <- c("x1_Sample", "x10_Sample", "x25_Sample", "x50_Sample", 
                           "x80_Sample", "x100_Sample")

# Creating grid of histograms
histGrid(Samples1000)

```


###Question 5 `diamonds$carat`

```{r}

# Adding Mins Samples1000 to SampleStats
SampleStats[5,] <- c(min(Samples1000[,1]), max(Samples1000[,2]), min(Samples1000[,3]), 
                     min(Samples1000[,4]), min(Samples1000[,5]), min(Samples1000[,6]))

# Adding Maxes Samples1000 to SampleStats
SampleStats[6,] <- c(max(Samples1000[,1]), max(Samples1000[,2]), max(Samples1000[,3]), 
                     max(Samples1000[,4]), max(Samples1000[,5]), max(Samples1000[,6]))

# Adding Means Samples1000 to SampleStats
SampleStats[7,] <- c(mean(Samples1000[,1]), mean(Samples1000[,2]), mean(Samples1000[,3]), 
                     mean(Samples1000[,4]), mean(Samples1000[,5]), mean(Samples1000[,6]))

# Adding variances Samples1000 to SampleStats
SampleStats[8,] <- c(var(Samples1000[,1]), var(Samples1000[,2]), var(Samples1000[,3]), 
                     var(Samples1000[,4]), var(Samples1000[,5]), var(Samples1000[,6]))

# Updating row names for newly added columns
rownames(SampleStats)[5:8] <- c("ResampleMin", "ResampleMax", "ResampleMean", "ResampleVariance")

# New SampleStats dataframe - transposed with new Resample rows added
SampleStats
```


###Question 6 `diamonds$carat`
```{r}
# Creating a histogram
sampleMeans <- data.frame(t(SampleStats["SampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=diamonds, aes(x=carat, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =sampleMeans,(aes(xintercept = SampleMean, colour = "red" )))
print(hist)
```


###Question 7 `diamonds$carat`
```{r}
# Creating a histogram
resampleMeans <- data.frame(t(SampleStats["ResampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=diamonds, aes(x=carat, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =resampleMeans,(aes(xintercept = ResampleMean, colour = "red" )))
print(hist)
```




####Question 1 `faithful$eruptions`

```{r}

# To get random samples, using the random samples function from class. 

# Calculating sample sizes for 1%, 10%, 25%, 50%, 80% samples
n1 <- as.integer(.01 * nrow(faithful))
n10 <- as.integer(.10 * nrow(faithful))
n25 <- as.integer(.25 * nrow(faithful))
n50 <- as.integer(.5 * nrow(faithful))
n80 <- as.integer(.8 * nrow(faithful))

# Getting random samples
abSample1 <- randomSample(faithful, n1, FALSE)
abSample10 <- randomSample(faithful, n10, FALSE)
abSample25 <- randomSample(faithful, n25, FALSE)
abSample50 <- randomSample(faithful, n50, FALSE)
abSample80 <- randomSample(faithful, n80, FALSE)

# Creating dataframes with min, max, mean, and variance for each sample 
abSample1 <- data.frame(c( min(abSample1$eruptions), max(abSample1$eruptions), mean(abSample1$eruptions), var(abSample1$eruptions) ))
abSample10 <- data.frame(c( min(abSample10$eruptions), max(abSample10$eruptions), mean(abSample10$eruptions), var(abSample10$eruptions) ))
abSample25 <- data.frame(c( min(abSample25$eruptions), max(abSample25$eruptions), mean(abSample25$eruptions), var(abSample25$eruptions) ))
abSample50 <- data.frame(c( min(abSample50$eruptions), max(abSample50$eruptions), mean(abSample50$eruptions), var(abSample50$eruptions) ))
abSample80 <- data.frame(c( min(abSample80$eruptions), max(abSample80$eruptions), mean(abSample80$eruptions), var(abSample80$eruptions) ))
abSample100 <- data.frame(c( min(faithful$eruptions), max(faithful$eruptions), mean(faithful$eruptions), var(faithful$eruptions) ))

# Combining dataframes of summary statistics from samples
SampleStats <- cbind(abSample1, abSample10, abSample25, abSample50, abSample80, abSample100)

# Naming rows and columns of SampleStats dataframe
colnames(SampleStats) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample", "100% Sample")
rownames(SampleStats) <- c("SampleMin", "SampleMax", "SampleMean", "SampleVariance")

# 6 columns and 4 fours, just like the question asks.
str(SampleStats)

```


####Question 2 `faithful$eruptions`

```{r}

# Running function of 
histGrid(SampleStats, transpose=TRUE)

```


####Question 3 `faithful$eruptions`

```{r}

# Creating an empty dataframe that will be used to store 1000 samples
Samples1000 <- data.frame("1% Sample" = 0, "10% Sample" = 0, "25% Sample" = 0, "50% Sample" = 0 , "80% Sample" = 0)

# Starting a for loop that iterates 1000 times
for(i in 1:1000) {
  
  # Taking random sample using sample size determined above 
  # Getting random samples
  abSample1 <- randomSample(faithful, n1, FALSE)
  abSample10 <- randomSample(faithful, n10, FALSE)
  abSample25 <- randomSample(faithful, n25, FALSE)
  abSample50 <- randomSample(faithful, n50, FALSE)
  abSample80 <- randomSample(faithful, n80, FALSE)
  
  # Adding sample to the empty dataset
  Samples1000[i,]<- c(mean(abSample1$eruptions),
                      mean(abSample10$eruptions),
                      mean(abSample25$eruptions),
                      mean(abSample50$eruptions),
                      mean(abSample80$eruptions))
  
}

# Now the dataframe has four columns coressponding to each the four calculated sample statistcs and 1000 rows for 1000 samples.
# The 1%, 10%, 25%, 50%, and 80% sample sizes each make up 1/5 of the dataframe
str(Samples1000)

```


###Question 4 `faithful$eruptions`

```{r}

# Adding the 100% sample to the dataframe
Samples1000[,6] <- abSample100[3,]

# Naming rows and columns of Samples1000 dataframe
colnames(Samples1000) <- c("x1_Sample", "x10_Sample", "x25_Sample", "x50_Sample", 
                           "x80_Sample", "x100_Sample")

# Creating grid of histograms
histGrid(Samples1000)

```


###Question 5 `faithful$eruptions`

```{r}

# Adding Mins Samples1000 to SampleStats
SampleStats[5,] <- c(min(Samples1000[,1]), max(Samples1000[,2]), min(Samples1000[,3]), 
                     min(Samples1000[,4]), min(Samples1000[,5]), min(Samples1000[,6]))

# Adding Maxes Samples1000 to SampleStats
SampleStats[6,] <- c(max(Samples1000[,1]), max(Samples1000[,2]), max(Samples1000[,3]), 
                     max(Samples1000[,4]), max(Samples1000[,5]), max(Samples1000[,6]))

# Adding Means Samples1000 to SampleStats
SampleStats[7,] <- c(mean(Samples1000[,1]), mean(Samples1000[,2]), mean(Samples1000[,3]), 
                     mean(Samples1000[,4]), mean(Samples1000[,5]), mean(Samples1000[,6]))

# Adding variances Samples1000 to SampleStats
SampleStats[8,] <- c(var(Samples1000[,1]), var(Samples1000[,2]), var(Samples1000[,3]), 
                     var(Samples1000[,4]), var(Samples1000[,5]), var(Samples1000[,6]))

# Updating row names for newly added columns
rownames(SampleStats)[5:8] <- c("ResampleMin", "ResampleMax", "ResampleMean", "ResampleVariance")

# New SampleStats dataframe - transposed with new Resample rows added
SampleStats
```


###Question 6 `faithful$eruptions`
```{r}
# Creating a histogram
sampleMeans <- data.frame(t(SampleStats["SampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=faithful, aes(x=eruptions, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =sampleMeans,(aes(xintercept = SampleMean, colour = "red" )))
print(hist)
```


###Question 7 `faithful$eruptions`
```{r}
# Creating a histogram
resampleMeans <- data.frame(t(SampleStats["ResampleMean",]))

hist <- ggplot() 
    # Updating plot to be histogram of density of distribution of statistic
    hist <- hist + geom_histogram(data=faithful, aes(x=eruptions, y = ..density..), binwidth=0.01)
    hist <- hist + geom_vline(data =resampleMeans,(aes(xintercept = ResampleMean, colour = "red" )))
print(hist)
```



##Question 8 --END


###Question 9

For all three data sets, we see that larger sample sizes lead to better estimates. Larger sample size and greater number of samples both lead to predicted means closer to the actual mean. The distribution of means from the sampleMeans column varies since there is only 1 sample at a given sample size. For all three of the datasets, the Resample Mean alines much closer to the actual mean and the variance of the distribution of means is much lower. 