---
title: "TEST1_R"
author: "Yogi Patel"
output: html_document
---


Importing data
```{r}
require(ggplot2)
require(grid)
require(plyr)
data(diamonds)
```

### Question One
Updating diamonds dataframe to include column with logical equivalent of the ratio of variable `vs` in dataframe `mtcars`.
```{r}
# Determining the ratio of 1's and 0's in mtcars$vs
x <- count(mtcars, 'vs')

# Calculating ratio of 1's in dataframe
propOnes <- x$freq[2]/(x$freq[2]+x$freq[1])
propOnes

# Determining number of rows in diamonds dataframe
numObs <- nrow(diamonds)
numObs

# Calculating number of Trues's needed in diamonds dataframe
numTrue <- propOnes * numObs
numTrue

########################################################################
# SIDE BAR -- In the email Gerhard sent for the interpretation of question one, the student reported that the number of Trues would be 23,402. But it looks like he reported the number of rows in Diamonds as 53,490, NOT 53,940 (the actual number).

# Just checking to see if I get the same number of Trues as reported in the email using the incorrect number of rows. 

propOnes * 53490

# Yes, using the 53490 get's the same number of Trues as in the email. I'll continue using the correct number of rows.

########################################################################

# Getting a whole number value for numTrue
numTrue <- round(numTrue,0)

# Adding a column, vs, to diamonds dataframe. Making all values false.
diamonds$vs <- FALSE
head(diamonds)

# Updating diamonds$vs column with correct number of True dataframes
diamonds$vs[0:numTrue] <- TRUE
head(diamonds)

# Confirming ratio of T/F is about equal (we had to round) in mtCars and diamonds

# Getting ratio of T/F's in diamonds$vs
y <- count(diamonds, 'vs')
propTrues <- y$freq[2]/(y$freq[2]+y$freq[1])

# Rounding proportions for comparison
propTrues <- round(propTrues, 3)
propOnes <- round(propOnes, 3)

# Checking if equal
propTrues == propOnes

# Finally, confirming that diamonds$vs is a logical vavriable
typeof(diamonds$vs)
```

The proportion of True to False in `diamonds$vs` is equal to the proportion of 1 to 0 in `mtcars$vs`. Also, we have confirmed that `diamonds$vs` is a logical variable. 




### Question Two

Creating function `getFactLogic` that will take one input parameter `df`, a dataframe. The function will return a new data frame including only the factor and logical columns in the dataframe. 

```{r}
# Creating function getFactLogic
getFactLogic <- function(df) {
  
  # Creating an empty dataframe to be populated with factor/logical variables
  # Number of rows is set to match input dataframe
  newdf <- data.frame(matrix(0, ncol = 0, nrow = nrow(df)))
  
  # Looping through each column to determine if they are factor/logical
  for (i in names(df)) {
    # Checking if variable is a factor variable
    if (is.factor(df[[i]])){
      # If it is a factor variable, add it to the new dataframe
      newdf[[i]] <- df[[i]]
    }
    # Checking if variable is a logical variable
    else if (is.logical(df[[i]])){
      # If it is a logical variable, add it to the new dataframe
      newdf[[i]] <- df[[i]]
    }
  }
  
  # Checking if any columns were added to the new dataframe.
  if (ncol(newdf) == 0){
    # If no columns were added, we do not want to return an empty dataframe.
    # Create error and exit function
    stop("No Logical or Factor variables found")
  }
  # If one or more columns are in dataframe, return new dataframe
  else { 
      return (newdf) 
      }
}

# Checking if factor/logical variables in diamonds are returned
y <- getFactLogic(diamonds)
head(y)
str(y)

# Checking if error works (mtcars does not contain factor/logical variables)
# Commented out so rest of document can run.
# x <- getFactLogic(mtcars)
```


### Question Three

Plotting a bar graph for every factor and logical variable.

First, getting a new data frame `diamondsPlot` with only factor and logical variables.
```{r}
diamondsPlot <- getFactLogic(diamonds)
head(diamondsPlot)
```

Now looping through columns in `diamondsPlot` to graph each one. 
```{r}
# Iterating through the columns in diamondsPlot
for (i in names(diamondsPlot)) {
# Creating a frequancy table for the column
x <- count(diamondsPlot, i)
# Creating a bar plot of the frequency
barplot(x$'freq', # plot using Frequency column
        main = i, # Title set to name of variable
        names.arg= x[[i]], # X lables set to values
        ylab = "Frequency") # Y axis name
}
```



### Question Four

Creating a list `freq` of frequency tables for factor/logical variables
```{r}
# Will continue to use diamondsPlot
freq <- list()
# Iterating through the columns in diamondsPlot
for (i in names(diamondsPlot)){
   # Generating frequency table and adding it to freq.
   freq[[i]] <- table(diamondsPlot[[i]])
}

freq
freq[1]